---
title: 'Exercise'
# author: TCornulier
output: 
  html_document: 
    code_folding: hide
    toc: false
editor_options: 
  chunk_output_type: console
---


```{r setup, echo=FALSE, purl = FALSE}
knitr::opts_chunk$set(echo=TRUE, message = FALSE, warning = FALSE, eval = FALSE, cache = FALSE)

SOLUTIONS<- FALSE
```

\  

# Exercise: Poisson GLM - predicting species richness

\  

The research question underpinning the data collection was to describe the relationships between the number of species (`Species`) present in a plot of land, and how this number was determined by the overall plant biomass (`Biomass`, i.e. what was the overall weight of plants in the plot of land), and the pH of the soil (`pH`, categorised into "High", "Medium" and "Low"). The researchers were interested in merely describing the relationships between species and the two explanatory variables. In total, the researchers collected data from 90 plots of land ($n = 90$), where each observation is from a single plot.

From this information alone, we are already able to determine a suitable distribution to use when fitting the model (the fact that this practical is called Poisson GLM may also offer a cryptic hint...). The relevant information comes from the description of `Species` clearly indicating that this variable is a count, which implies 1) the minimum value is 0 [we cannot have -1 species], and 2) the response variable will be measured in integers [we cannot count 3.14 plants]. With this information alone, we can already deduce the *stochastic* element of our model should be adequately described by specifying a $Poisson$ distribution (but note that there are other distributions we could use).

\  

1\. As in previous exercises, either create a new R script (perhaps call it GLM_Poisson) or continue with your previous R script in your RStudio Project. Again, make sure you include any metadata you feel is appropriate (title, description of task, date of creation etc) and  don't forget to comment out your metadata with a `#` at the beginning of the line. 

\  

2\. Import the data file 'species.txt' into R and take a look at the structure of this dataframe. Given you have never seen this data before, it's really important that you familiarise yourself with any nuances. To help with this, carry out an initial data exploration (using any methods you think will help you get a sense of the data, e.g. `plot()`, `pairs()`, `coplot()`, amongst many other options).

While doing this, ask yourself:

* Do any of the variables need to be adjusted? (e.g. are factors recognised as such?)
* Do any factors need to be "re-levelled", such that it is read "Low", "Medium "High" (or any order we may prefer)?
  + (Hint: check `?factor` and look at the `levels` argument)
* Are there any relationships you can already see by eye alone?
* Are there any imbalances in any of the explanatory variables?
* Are there any observations that seem like they may be a data entry mistake?

If using `pairs()` to create a plot of the variables of interest, rather than creating a plot with every single variable in our dataset, we may prefer to restrict the plot to the variables we are actually interested in. An effective way of doing this is to store the names of the variables of interest in a vector `VOI<- c("Var1", "Var2", ...)` and then use the list of variable names to subset the variables that get plotted (e.g. `pairs(Mydata[, VOI])`)


\  

```{r Q2, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}
sp<- read.table(file= "./data/species.txt", header= TRUE)

# Check the structure of the data to see if we need to correct anything
str(sp)

# Correct pH so it is read as a factor, reordered such that "low" is the reference level
sp$pH<- factor(sp$pH, levels = c("low", "mid", "high"))

# Get simple descriptives of the data (e.g. what is the range for each variable?)
summary(sp)

# make a list of the variables of interest, for convenience:
VOI<- c("Species", "Biomass", "pH")
pairs(sp[, VOI])
# Negative relationship between Species and Biomass?
# Positive relationship between Species and pH? 
# Biomass tends to increase with pH, which could generate
# some collinearity between these explanatory variables in a model.
# but still plenty of variation in Biomass within each pH, 
# so hopefully this won't be an issue.

coplot(Species ~ Biomass | pH, data = sp)
# the relationships looks consistently negative with biomass across the pH levels
```

\  

3\. To warm up, let's start with a simple Poisson GLM (using `glm`) where we assume that the number `Species` counted in a plot is a function of how much `Biomass` there was. 

As a reminder, for this model the stochastic element is specified by using the $Poisson$ distribution, which controls *how* randomness is allowed to work for this data (i.e. even if a given amount of `Biomass` would predict 5 species, we wouldn't expect zero variation away from this but neither would we think a prediction of one billion species is sensible - the $Poisson$ distribution helps ensure that any randomness is appropriate for the type of data we're working with). The deterministic element of this model will be "saying" that at least part of what explains the number of species is determined by how much biomass there is in a given plot. 

* Hints:
  + We use `family =` to specify the distribution in a `glm()`
    - "Family" is just an alternative way to refer to a distribution.
  + What is the default link function used by Poisson GLMs?
    - How do we specify it?
    - Do we need to specify it?

\  

```{r Q3, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}
sp.glm1 <- glm(Species ~ Biomass, family = poisson(link = "log"), data = sp)
```

\  

4\. Obtain summaries of the model output using the `summary()` function. Make sure you understand the mathematical and biological interpretation of the model, by writing down the complete model on paper.

For example, an intercept-only $Poisson$ GLM would be:

$$
y_i \sim Poisson(\lambda_i)\\
log(\lambda_i) = \beta_0
$$

where, $y_i$ is the value of the response variable for the $i^{th}$ observation (e.g. if $i = 1$, then this is the first row of our dataset), assumed to follow a $Poisson$ distribution with rate $/lambda$, regressed on the $log$ link as a function of an intercept ($\beta_0$) only.

How would you extend this, given we've included `Biomass` in our model? If you've run the model, you will have an estimate for your parameters ($\beta_0$ and any other you may have included), so we can replace these temporary place holder labels (e.g. $\beta_0$) with the estimated values.

\  

```{r Q4, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}
summary(sp.glm1)

# Model description:
# Species_i ~ Poisson(mu_i)
# log(mu_i) = 3.18 - 0.064*Biomass_i

```

\  

5\. According to this model, what would be the predicted value (by hand) on the link scale (i.e. $log(\lambda)$) when `Biomass` is equal to 5? What would this become on the response scale, i.e. the number of species?

\  

```{r Q5, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}
# On the link scale:
3.18 - 0.064*5 # 2.86
# On the response scale (species count):
exp(3.18 - 0.064*5) # 17.46
```

\  

6\. We've now run a Poisson GLM where `Species` is explained by `Biomass`, but remember that our initial research question sought to describe the relationships between `Species` with both `Biomass` and `pH`. Before running this model, try to write down the equation for this new version of the model (you can always go back and watch any relevant sections of the Poisson lecture). Once you've given that a shot, run the `glm()` which includes both `Biomass` and `pH` as main effects only (we won't be doing any interactions yet).

\  

```{r Q6, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}
sp.glm2 <- glm(Species ~ Biomass + pH, family = poisson(link = "log"), data = sp)
```

\  

7\. Obtain summaries of the model output using the `summary()` and the ANOVA of the model. 

* Which of the `drop1()` or `anova()` functions would you choose to use if you wanted to
  + look at deviance components?
  + explain the predictions?
  + test hypotheses?
  + simplify the model?
* Is the effect of the interaction significant? 

\  

```{r Q7, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}
summary(sp.glm2)
anova(sp.glm2, test= "Chisq")
# DEVIANCE COMPONENTS:
# anova is useful for getting an idea of the relative contributions
# of each term to explaining the overall variability in the data,
# using the deviance. Beware that order does matter for all but the last term!

# EXPLAINING PREDICTIONS:
# to explain patterns in the data, we need to look at the coefficient
# estimates from the summary of the model.

# TESTING HYPOTHESES
# To test hyoptheses on the coefficient value being different from zero,
# look at the Z-test in the summary table

# To test hyoptheses on a predictor explaining a significant porportion of 
# variation in the data, look at the Chi-sq-test in the anova table

# MODEL SELECTION (SIMPLIFICATION):
# In general 'drop1' would be the method of choice if our purpose is 
# to perform model simplification.
# 'anova' would provide the same answer in this case, since there are
# no other interactions involved (and interactions are always last). 

drop1(sp.glm2, test= "Chisq")

# the ANOVA table provides a global test across pH levels. 
# However the significance of the interactions is also unambiguous
# from the summary table, in this case, although the null hypothesis
# is different (coefficient different from zero in the latter 
# vs. significant proportion of variation explained in the former).

```

\  

8\. Looking at the model summary, make sure you understand the individual components, null hypotheses, and the mathematical and biological interpretation of the different coefficients (i.e. would you be able to reconstruct and to use the model formula to make predictions? In doubt, try it and seek assistance!). 

* How do you interpret the model terms? 
  + what does the `Intercept` represent, and which categories does it apply to?
  + what does the `Biomass` effect represent, and which category does it apply to?
  + what does the `pHmid` effect represent?
  + what does the `Biomass:pHmid` effect represent?

\  

```{r Q8, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}

# "(Intercept)" is the predicted value on the link (log) scale for
# Biomass = 0, and pH= "low"

# "Biomass" is the slope of Biomass for the low pH category.
# It is negative, so assumes a linear decrease (on the log scale)

# "pHmid" is the estimated difference (on the log scale) between
# the "pHmid" and the reference level, "pHlow".
# Positive means higher on average than pHlow.

# "Biomass:pHmid" is the difference between the slope of Biomass for pHmid
# compared to the slope of Biomass for the reference level, "pHlow".
# Positive means that the decrease is slower in "pHmid" (although this is
# for the linear relationships on the log scale, but it may not look
# the same on the response scale - see the graphical interpretation below)

# A mathematical description of the model
# (more or less how I would present it in the methods section of a paper):
# Species ~ Poisson(mu)
# log(mu) = 2.95 - 0.26 * Biomass
#    + 0.48 * pHmid
#    + 0.82 * pHhigh
#    + 0.12 * Biomass * pHmid
#    + 0.16 * Biomass * pHhigh

# The coefficients are on the log scale,
# so cannot be interpreted directly as counts. 
# They are interpreted as changes in log-counts:

# For a biomass of zero, the log of the number of species at medium pH
# is increased by 0.48 compared to the low pH.
# This is equivalent to saying that the counts increase by exp(0.48),
# hence that they are multiplied by 1.62.

# The summary gives a residual deviance of 83 on 84 degrees of freedom,
# so the ratio is about 1, hence no indication of under- or over-dispersion

```

\  

9\. Any indication of overdispersion? (Hint: check residual deviance and degrees of freedom)

\  

```{r Q9, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}

# "(Intercept)" is the predicted value on the link (log) scale for
# Biomass = 0, and pH= "low"

# "Biomass" is the slope of Biomass for the low pH category.
# It is negative, so assumes a linear decrease (on the log scale)

# "pHmid" is the estimated difference (on the log scale) between
# the "pHmid" and the reference level, "pHlow".
# Positive means higher on average than pHlow.

# "Biomass:pHmid" is the difference between the slope of Biomass for pHmid
# compared to the slope of Biomass for the reference level, "pHlow".
# Positive means that the decrease is slower in "pHmid" (although this is
# for the linear relationships on the log scale, but it may not look
# the same on the response scale - see the graphical interpretation below)

# A mathematical description of the model
# (more or less how I would present it in the methods section of a paper):
# Species ~ Poisson(mu)
# log(mu) = 2.95 - 0.26 * Biomass
#    + 0.48 * pHmid
#    + 0.82 * pHhigh
#    + 0.12 * Biomass * pHmid
#    + 0.16 * Biomass * pHhigh

# The coefficients are on the log scale,
# so cannot be interpreted directly as counts. 
# They are interpreted as changes in log-counts:

# For a biomass of zero, the log of the number of species at medium pH
# is increased by 0.48 compared to the low pH.
# This is equivalent to saying that the counts increase by exp(0.48),
# hence that they are multiplied by 1.62.

# The summary gives a residual deviance of 83 on 84 degrees of freedom,
# so the ratio is about 1, hence no indication of under- or over-dispersion

```

\  

10\. [Optional] Check for collinearity using the `vif()` function in the `car` package. What do you think?

\  

```{r Q10, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}
library(car)
vif(sp.glm2)

# GVIF is high for all the terms.
# This is expected though, because all terms are part of an interaction. 
# Any change in one coefficient will have a strong impact on the others
# This might be exacerbated by the pre-existing correlation between 
# the predictors. But all components are very clearly needed here.

```

\  

11\. Validate the model using the standard residuals diagnostic plots

\  

```{r Q11, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE, fig.show= ifelse(SOLUTIONS, "asis", "hide")}
par(mfrow= c(2, 2))
plot(sp.glm2)

# nothing particularly bad... 
# the variance of the standardized residuals ("Pearson") tends to decrease 
# slightly, when it should be constant (underdispersion).
# a few observations poorly predicted (18, 20) and slight
# overestimation of the smaller values (left end of top-left graph), 
# just to be picky.
# There is no expectation that the residuals should be normally distributed 
# (thus QQplot not particularly useful)

# we can produce more residual plots if desired:
# extract various residuals and plot
res.p<- resid(sp.glm2, type= "pearson")
res.d<- resid(sp.glm2, type= "deviance")

fit<- fitted(sp.glm2) # on the response scale

plot(res.p, res.d); abline(0,1) # deviance against Std Pearson residuals
# quite similar in this model (no overdispersion, moderate underdispersion)
plot(fit, res.d, ylab= "Deviance residuals", xlab= "Fitted values (response scale)")
abline(h= 0, col= 2, lty= 2)

# residuals against the predictors:
plot(sp$pH, res.p, ylab= "Std Pearson residuals", xlab= "pH")
abline(h= 0, col= 2, lty= 2)
# low pH plots are more variable

plot(sp$Biomass, res.p, ylab= "Std Pearson residuals", xlab= "Biomass")
abline(h= 0, col= 2, lty= 2)
# Species richness over-predicted for highest and lowest Biomass plots
# Few cases: could be by chance?

```

\  

12\. Let's plot the model predictions over the raw data, using `predict()` with the argument `type= "response"` to obtain the fitted values on the original (response) scale. Plot a fitted line for the relationship between number of species and biomass for each level of pH. For the sake of this exercise and to better illustrate the assumptions of the model, use the whole range of biomass(0 to 10) for every pH level. Why are the lines not straight?

\  

```{r Q12, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE, fig.height=8, fig.show= ifelse(SOLUTIONS, "asis", "hide")}
par(mfrow= c(1, 1))
# create a sequence of increasing Biomass
Biomass.seq<- seq(from= min(sp$Biomass), to= max(sp$Biomass), l= 25)
# predict for a range of biomass values and a low pH
MyData1<- data.frame(Biomass= Biomass.seq, pH= "low")
# predict for a range of biomass values and a mid pH
MyData2<- data.frame(Biomass= Biomass.seq, pH= "mid")
# predict for a range of biomass values and a high pH
MyData3<- data.frame(Biomass= Biomass.seq, pH= "high")

P.low<- predict(sp.glm2, newdata= MyData1, type= "response")
P.mid<- predict(sp.glm2, newdata= MyData2, type= "response")
P.high<- predict(sp.glm2, newdata= MyData3, type= "response")

plot(sp$Biomass, sp$Species, col= sp$pH, xlab= "Biomass", ylab= "N species")
# when pH is converted to numeric, 
# you get 1 for "low", 2 for "med" and 3 for "high"
# color 1 means black in R
# color 2 means red in R
# color 3 means green in R

lines(MyData1$Biomass, P.low, lty= 1, col= 1)
lines(MyData2$Biomass, P.mid, lty= 1, col= 2)
lines(MyData3$Biomass, P.high, lty= 1, col= 3)

legend("topright", 
 legend= c("high", "mid", "low"), 
 col= c(3:1), 
 lty= c(1, 1, 1), 
 lwd= c(1, 1, 1))
```

\  

13\. Looking at the data and model fits, any idea why the expected species richness for the largest values of biomass tends to be biased high? How satisfied are you with the model? Have we learned anything from it, about species diversity in relation to pH and biomass? 

\  

```{r Q13, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE}
# The data show a very steep decline in species richness towards zero, 
# as biomass increases.

# The log-link allows to transform the straight lines (on the link scale) 
# into curves (on the response scale), while ensuring that predictions
# are always positive. But the log link doesn't allow enough flexibility
# for the decline to happen as fast as it does in the data
# while remaining positive for any value of biomass.

# The model (a straight line for each pH) is too simple to capture this 
# level of detail in the data, and hence is biased.

# However it does a pretty good job overall, and can be trusted to 
# make robust inference about the general trends.

# Note that the interaction is not obvious on the response scale,
# as the curves look quite "parallel". It is indeed one of the 
# peculiarities of transformations (like the log), that parallel lines
# on one scale will not be so on the other scale.
# Caution is therefore required in interpreting effects from the 
# summary of the model directly, and plotting on the response scale 
# is highly recommended before drawing any conclusion!

```

\  

14\. (Optional) Use `predict` again, but this time,
* obtain the fitted values on the scale of the linear predictor `type= "link"`
* only predict for the range of observed biomass for each pH level. 
Plot again and compare with the previous graph: what is happening? How would you back-transform these values predicted on the link scale to plot them on the response scale again?

\  

```{r Q14, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE, fig.height=8, fig.show= ifelse(SOLUTIONS, "asis", "hide")}
Biomass.seq1<- seq(from= min(sp$Biomass[sp$pH== "low"]), 
                    to= max(sp$Biomass[sp$pH== "low"]), 
                    l= 25)
Biomass.seq2<- seq(from= min(sp$Biomass[sp$pH== "mid"]), 
                    to= max(sp$Biomass[sp$pH== "mid"]), 
                    l= 25)
Biomass.seq3<- seq(from= min(sp$Biomass[sp$pH== "high"]), 
                    to= max(sp$Biomass[sp$pH== "high"]), 
                    l= 25)

# predict for a range of biomass values and a low pH
MyData1<- data.frame(Biomass= Biomass.seq1, pH= "low")
# predict for a range of biomass values and a mid pH
MyData2<- data.frame(Biomass= Biomass.seq2, pH= "mid")
# predict for a range of biomass values and a high pH
MyData3<- data.frame(Biomass= Biomass.seq3, pH= "high")

P.low<- predict(sp.glm2, newdata= MyData1, type= "link")
P.mid<- predict(sp.glm2, newdata= MyData2, type= "link")
P.high<- predict(sp.glm2, newdata= MyData3, type= "link")

plot(sp$Biomass, sp$Species, col= sp$pH, xlab= "Biomass", ylab= "Number of species")

lines(MyData1$Biomass, P.low, lty= 1, col= 1)
lines(MyData2$Biomass, P.mid, lty= 1, col= 2)
lines(MyData3$Biomass, P.high, lty= 1, col= 3)
# lines are straight and appear in the wrong place!
# Because they are on the link scale (observations are on the response scale)

# Now, re-plot after back-transforming the predictions to make them on the response scale
plot(sp$Biomass, sp$Species, col= sp$pH, xlab= "Biomass", ylab= "Number of species")

# (note the 'exp')
lines(MyData1$Biomass, exp(P.low ), lty= 1, col= 1)
lines(MyData2$Biomass, exp(P.mid ), lty= 1, col= 2)
lines(MyData3$Biomass, exp(P.high), lty= 1, col= 3)
# Better?

legend("topright", 
 legend= c("high", "mid", "low"), 
 col= c(3:1), 
 lty= c(1, 1, 1), 
 lwd= c(1, 1, 1))
```

\  

15\. (Optional but recommended) Use `predict` again, but this time obtain the fitted values and their standard errors on the scale of the linear predictor. From these, calculate confidence intervals for the fitted values, and plot them together with the data, after back-transforming the fitted values and intervals on the response scale.

* Suggested approach:
  + plot the raw data (one colour per pH)
  + create a `data.frame` called `X` containing the data to predict for: as sequence of increasing Biomass and the pH of your choice
  + use `predict()` with the appropriate options `type= "link", se.fit= TRUE` to obtain the fitted values on the link scale and for being able to calculate the confidence intervals later. Store in object `Z`.
  + plot fitted values, extracted using `Z$fit`, against the Biomass sequence. Do not forget to back-transform on the response scale.
  + plot the upper bound of the 95% CI (fitted values + 1.96*se), extracted using `Z$fit` and `Z$se.fit`, against the Biomass sequence. Do not forget to back-transform on the response scale.
  + plot the lower bound of the 95% CI (fitted values - 1.96*se), extracted using `Z$fit` and `Z$se.fit`, against the Biomass sequence. Do not forget to back-transform on the response scale.
  + repeat for other pH values.

\  

```{r Q15, eval=TRUE, echo=SOLUTIONS, results=SOLUTIONS, collapse=TRUE, fig.height=8, fig.show= ifelse(SOLUTIONS, "asis", "hide")}
P.low<- predict(sp.glm2, newdata= MyData1, type= "link", se.fit= T)
P.mid<- predict(sp.glm2, newdata= MyData2, type= "link", se.fit= T)
P.high<- predict(sp.glm2, newdata= MyData3, type= "link", se.fit= T)

plot(sp$Biomass, sp$Species, col= sp$pH, xlab= "Biomass", ylab= "Number of species")

# back-transform the predictions to make them on the response scale
# (note the 'exp')
lines(MyData1$Biomass, exp(P.low$fit ), lty= 1, col= 1)
lines(MyData2$Biomass, exp(P.mid$fit ), lty= 1, col= 2)
lines(MyData3$Biomass, exp(P.high$fit), lty= 1, col= 3)

legend("topright", 
 legend= c("high", "mid", "low"), 
 col= c(3:1), 
 lty= c(1, 1, 1), 
 lwd= c(1, 1, 1))

# (optional: overlay predictions for the full range of biomass, in dashed lines)
# create a sequence of increasing Biomass again
Biomass.seq<- seq(from= min(sp$Biomass), to= max(sp$Biomass), l= 25)
MyData1<- data.frame(Biomass= Biomass.seq, pH= "low")
MyData2<- data.frame(Biomass= Biomass.seq, pH= "mid")
MyData3<- data.frame(Biomass= Biomass.seq, pH= "high")

P.low<- predict(sp.glm2, newdata= MyData1, type= "link", se.fit= T)
P.mid<- predict(sp.glm2, newdata= MyData2, type= "link", se.fit= T)
P.high<- predict(sp.glm2, newdata= MyData3, type= "link", se.fit= T)

# back-transform the predictions to make them on the response scale
# (note the 'exp' and new line type 'lty= 2')
lines(MyData1$Biomass, exp(P.low$fit ), lty= 2, col= 1)
lines(MyData2$Biomass, exp(P.mid$fit ), lty= 2, col= 2)
lines(MyData3$Biomass, exp(P.high$fit), lty= 2, col= 3)


# Now add the lower bound of the 95% CI
lines(MyData1$Biomass, exp(P.low$fit  - 1.96*P.low$se.fit ), lty= 3, col= 1)
lines(MyData2$Biomass, exp(P.mid$fit  - 1.96*P.mid$se.fit ), lty= 3, col= 2)
lines(MyData3$Biomass, exp(P.high$fit - 1.96*P.high$se.fit), lty= 3, col= 3)

# and the upper bound of the 95% CI
lines(MyData1$Biomass, exp(P.low$fit  + 1.96*P.low$se.fit ), lty= 3, col= 1)
lines(MyData2$Biomass, exp(P.mid$fit  + 1.96*P.mid$se.fit ), lty= 3, col= 2)
lines(MyData3$Biomass, exp(P.high$fit + 1.96*P.high$se.fit), lty= 3, col= 3)

# The confidence intervals demonstrate clearly the difference between all pH levels.
# Note the variable width, with more uncertainty at the extremes of the biomass data range.
# If you have used the full extent of biomass for the 'low' pH level like in my code here,
# you will also notice that the confidence interval shrinks again,
# as the predicted values approach zero. 
# This is the back-transformation doing its job, 
# forcing fitted values to remain positive!
```

\  
\  

**End of the Poisson GLM - predicting species richness exercise**

